---
title: "linear_models_1"
author: "Caleigh Dwyer"
date: "2023-11-09"
output: github_document
---

lecture notes

##linear models

Observe data for subjects 1 to n. Want to estimate the coefficients in the model.

y = b0 (the y intercept) + B1x1 (b1 is the effect of the predictor on the relationship between x and y)

Assumptinos: residuals have mean zero, constant variance, and are independent

estimate parameters using OLS

Outcome is continuous, predictors can be anything. Continuous predictors are added directly

categorical predictors require "dummy" or indicator variable
-For each non-reference group, a binary 0/1 variable indicating group membership for each subject is created and used in the model

B1 = (B0 + B1 + B2) - (B0 + B2)
  = E(y| age = 1, sex = male) - E(y|age = 0, sex = male)
  [E is the expected value]
  expected change in y for a one unit cahnge in age, keeping sex fixed

##Testing

For a single regression coefficient, you can construct a test statistic:

t = (B -b)/se(B)

for large samples, the test statistic has a standard normal distribution

to test multiple coefficients (i.e. categorical variable w/ multiple predictros) you can use an F test (ANOVA)

##Diagnostics

Many model assumptions (constant variance, model specification, etc.) can be examined using residuals
- look at overall distribution (centered? skewed? outliers?)
- look at residuals vs. predictors (any non-linearity? Trends? Non-constant residual variance?)

##generalized linear models

appropriate for non-continuous outcomes
common example is logistic regression:

logit(P(Y=1|x)/P(Y=0)|x) = b0 + b1x1...

b1 for a log regression is an odds ratio, so very interpretable. 

##linear models in R

lm for linear models

glm for generalized linear models

arguments include:
Formula: y ~ x1 + x2
Data

output is complex, and also kind of a mess, so we use the broom package. Broom package is used for a lot of different functions/commands

define data frame
fit model
tidy results


```{r}
library(tidyverse)
library(p8105.datasets)

set.seed(1)

```


```{r}
data("nyc_airbnb")

nyc_airbnb = 
  nyc_airbnb |> 
  mutate(stars = review_scores_location / 2) |> 
  rename(
    borough = neighbourhood_group,
    neighborhood = neighbourhood) |> 
  filter(borough != "Staten Island") |> 
  select(price, stars, borough, neighborhood, room_type)
```


##fit a model

```{r}
nyc_airbnb |> 
  ggplot(aes(x= stars, y = price, color = borough))+
  geom_point()
```


```{r}
fit = lm(price ~ stars + borough, data = nyc_airbnb)
```

let's look at the result...

```{r, include = FALSE}
fit

summary(fit)
summary(fit)$coef
coef(fit)
fitted.values(fit)
residuals(fit)
##we don't usually look at these things because it's a matrix and kinda messy
```


```{r}
broom::glance(fit)

##this formats as a tibble

broom::tidy(fit)

#this is an even better dataframe. will show you the estimate for each category (borough) compared to the reference group

broom::tidy(fit) |> 
  select(-std.error, -statistic) |> 
  mutate(
    term = str_replace(term, "borough", "Borough: ")
  ) |> 
  knitr::kable(digits = 3)

##this gives you a very clean table

```


## Be in control of factors
```{r}
nyc_airbnb =
  nyc_airbnb |> 
  mutate(
    borough = fct_infreq(borough),
    room_type = fct_infreq(room_type))
  )

nyc_airbnb |> 
  ggplot(aes(x= stars, y = price, color = borough))+
  geom_point()

##use factor reorder to organize borough by frequency of room type, this will make sure that the reference group used by the model is the one with the most observations. When you wanna start interpreting coefficients, it's important to use a logical/feasible reference group
```



```{r}
fit = lm(price ~ stars + borough, data = nyc_airbnb)

broom::tidy(fit)

broom::glance(fit)
```


##Diagnostics

```{r}
residuals(fit)

modelr::add_residuals(nyc_airbnb, fit)

##this shows how well your model fits

modelr::add_residuals(nyc_airbnb, fit) |> 
  ggplot(aes(x= borough, y = resid))+
  geom_violin()+
  ylim(-500, 1500)

##shows that manhattan and brooklyn have some really big residuals (due to outlying data)

nyc_airbnb |> 
  modelr::add_residuals(fit) |> 
  ggplot(aes(x=stars, y = resid))+ 
  geom_point()+
  facet_wrap(.~borough)

##shows wider distribution of residuals as you increase in star count in manhattan and brooklyn
```

##hypothesis tests

this does t test by default
```{r}
fit |> 
  broom::tidy()
```


what about significance of borough? You wanna use at anova (in the above,  you were only getting comparisons to the reference)

```{r}
fit_null = lm(price ~stars, data = nyc_airbnb)
fit_alt = lm(price ~stars + borough, data = nyc_airbnb)

anova(fit_null, fit_alt) |> 
  broom::tidy()

```


##nest data, fit models


this is pretty formal and also complex
```{r}
fit = lm(price ~ stars * borough + room_type * borough, data = nyc_airbnb)

broom::tidy(fit)

```

this is more exploratory but maybe easier to understand

```{r}
nyc_airbnb |> 
  nest(data = -borough) |> 
  mutate(
    models = map(.x = data, ~lm(price~stars, data =.x)),
    results = map(models, broom::tidy)
  ) |> 
  select(-data, -models) |> 
  unnest(results) |> 
  filter(term == "stars")

##shows you the relationship between borough and stars by borough (so not comparing across boroughs or in reference to each other)

nyc_airbnb |> 
  nest(data = -borough) |> 
  mutate(
    models = map(.x = data, ~lm(price~stars, data =.x)),
    results = map(models, broom::tidy)
  ) |> 
  select(-data, -models) |> 
  unnest(results) |> 
  filter(term != "(Intercept)") |> 
  select(borough, term, estimate) |> 
  pivot_wider(
    names_from = borough,
    values_from = estimate)
  )
```

let's nest even more

```{r}

nyc_airbnb |> 
  filter(borough == "Manhattan") |> 
  nest(data = -neighborhood)|> 
  mutate(
    models = map(.x = data, ~lm(price~stars, data =.x)),
    results = map(models, broom::tidy)
  ) |> 
  select(-data, -models) |> 
  unnest(results)

##shows the impact of stars on price by neighborhood

nyc_airbnb |> 
  filter(borough == "Manhattan") |> 
  nest(data = -neighborhood)|> 
  mutate(
    models = map(.x = data, ~lm(price~stars, data =.x)),
    results = map(models, broom::tidy)
  ) |> 
  select(-data, -models) |> 
  unnest(results) |> 
  filter(str_detect(term, "room_type")) |> 
  ggplot(aes(x = neighborhood, y = estimate))+
  geom_point() +
  facet_wrap(.~term)

##not sure why viz isn't working
```

